Gibbons-Hawking-York Boundary Term in Causal Sets

### 2014-09-18 paper_v2_backup_20140520.tex
initial commit

@@

\section{Introduction}

First and foremost one must draw attention to the fact that Michel Buck (on of the authors of this paper) is in fact a smelly son of bitch. Gerrrl, he so ugly he look like someone tried to put a fire out on his face with a fork!

In furthering causal set theory it is crucial that we understand the kinematics of the theory. The action of a given causal set is one such kinematical property that would be extremely useful to know and understand. Proposals for the action of a causal set are available [[REFERENCE]] and these hold analytically in some cases and numerically in many more. These cases being when the causal set is embeddable in some existing spacetime. One can then show that the action of the causal set then agrees with the Einstein-Hilbert action of the spacetime in some continuum limit. How this limiting procedure is carried out will be described in more detail below.

The EH action, however, is not the complete story in the continuum.

If one has a spacetime manifold with a boundary, such as Minkowski space but cut off at some finite time, then one needs to include the Gibbons-Hawking-York boundary term (GHYBT) in the Einstein-Hilbert action. This term is needed to keep the variational principle well defined. In this section we prove a conjecture for the form of this boundary term in causal set theory.

\section{The Claim}

Consider a sufficiently well-behaved spacetime $(M,g)$ and Cauchy surface $\Sigma$ in $M$. The causal past and future sets $M^\pm=J^\pm(\Sigma)$ form a partition of $M$ and $\partial M^\pm = \Sigma$. The Gibbons-Hawking-York boundary term for $M^\pm$ is in this case given by
$$
S_{GHY} = \pm\frac1{\kappa^2} \int_{\Sigma} d^{d-1}x \sqrt{h} K
$$
where $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$.

The extrinsic curvature of a spacelike hypersurface can be identified with the volume gradient across that hypersurface. On the causal set, the analogue of a spacelike hypersurface is a maximal anti-chain, i.e. a subset $A$ of $C$, no two elements of which are related. The intuitive analogue of the boundary term would then be the rate of change of the number of causal set elements below and above the antichain. We shall see that this intuitive idea indeed bears out.

Consider a causal set $\mathcal C$ obtained by a Poisson sprinkling of density $\rho$ into a spacetime $(M,g)$. A Cauchy surface $\Sigma\subset M$ induces a partition $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$ of the sprinkling, where $\mathcal C^\pm$ denotes the restriction of $\mathcal C$ to the points sprinkled to the causal future/past of $\Sigma$. Let us denote the number of maximal elements in $\mathcal C^-$ and the number of minimal elements in $\mathcal C^+$ by $N_{min}$ and $N_{max}$, respectively. We propose the following definition for the discrete Gibbons-Hawking-York boundary term for a causal set $\mathcal C$ in $d$ spacetime dimensions:
\be\label{GH_boundary_to_causet}
S^{(d)}_{GHY}[\mathcal C]=\rho^{\frac{2}{d}-1}c_{d}\left\langle N_{max}-N_{min}\right\rangle.
\ee
The constant $c_{d}$ only depends on the dimension and is given by

@@

To support this proposal we show that in the limit of infinite sprinkling density we obtain
\be
\lim_{\rho\rightarrow\infty}S^{(d)}_{GHY}[\mathcal C] = \frac1{\kappa^2} \int_{\Sigma} d^{d-1}x \sqrt{h} K.\label{eq:mainconjecture}
\ee
%MB1: (too vague) During the calculation approximations will be made. These approximations are justified on the grounds that if certain orders are retained throughout the calculation they will in fact vanish in the limit $\rho \rightarrow \infty$

\section{Poisson Sprinklings for $\langle N_{max}\rangle$ and $\langle N_{min}\rangle$}

@@

Both quantities will diverge, but if their difference $\langle N_{max}\rangle - \langle N_{min}\rangle = \langle N_{max} - N_{min}\rangle$ grows slower than or at order $\rho^{1-\frac2d}$, the proposed action~\eqref{GH_boundary_to_causet} will tend to a finite value in the continuum limit.  Figure \ref{fig:Nmin_Nmax} shows a 2-dimensional spacetime with an extrinsically curved spacelike surface (a line). The minimal and maximal points have been highlighted and the shaded regions correspond to those that  we wish to find the volumes of. This is a simple task here as the example is only 2D and the spacetime is flat.

\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{minmaxplot}
     \caption{Minimal and maximal points about a surface in 2D Minkowski space. The shaded regions illustrate the regions that one must calculate the volume of.}
     \label{fig:Nmin_Nmax}
\end{figure}

@@

\begin{align}\label{eq:nmax_and_eq:nmin}
\left\langle N_{max}\right\rangle & =\int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
h^{\frac{1}{2}}\left(1+
\frac{1}{2}\frac{\dot{h}}{h}t+O(t^2)\right)
 \rho\ e^{-\rho V(-t,\mathbf{x};0,\mathbf{x})}
\\
\left\langle N_{min}\right\rangle & =\int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
h^{\frac{1}{2}}\left(1+
\frac{1}{2}\frac{\dot{h}}{h}t+O(t^2)\right) \rho\ e^{-\rho V(0,\mathbf{x};t,\mathbf{x})}
\end{align}
where $h\equiv det\left(h_{ij}(0,\mathbf{x})\right)$ and $\dot{}\equiv \frac{\partial}{\partial t}$. The metric determinant has been in expanded in small $t$. The volume functions have been rewritten as follows: $V(x,\Sigma)=V(x,x_0)=V(-t,\mathbf{x};0,\mathbf{x})$ and $V(\Sigma,q)=V(x_0,x)=V(0,\mathbf{x};t,\mathbf{x})$

The only curvy cones that will contribute will have small volumes, as we are restricting ourselves close to the surface. This means that the volumes can be approximated by the volumes of less curvy cones, and it can be shown that higher order corrections, in the approximation scheme, vanish in the limit of $\rho \rightarrow \infty$. This will all be made more precise below. The approximation procedure will be outlined below for a cone to the future of $\Sigma$, as the process for the past cone is nearly identical. Riemann normal coordinates are integral to the following calculation so it is worth noting the relevant formulas before jumping in.

\section{Riemann Normal Coordinates}

One can always find Riemann normal coordinates (RNC), about a point $p$, by changing from coordinates $x^{\alpha}$ to coordinates $y^{\overline{\alpha}}$ such that in these new coordinates the metric and the Christoffel symbols at $p$ are that of flat space. These conditions  can be expressed as follows

\be\label{eq:RNCMetricTransAtPAndChris}
g_{\overline{\alpha} \overline{\beta}}(p)=\eta_{\overline{\alpha} \overline{\beta}}=A^{\alpha}_{\;\overline{\alpha}}A^{\beta}_{\;\overline{\beta}}g_{\alpha\beta}(p)\;,\;\;\;\;\Gamma^{\overline{\alpha}}_{\;\overline{\beta}\overline{\gamma}}(p)=0
\ee
The $A^{\alpha}_{\;\overline{\alpha}}$ govern the coordinate transformation to linear order, but $O(x^2)$ corrections may be required. To second order the coordinate transformation is given by

\be\label{eq:RNCtotaltrans}
y^{\overline{\alpha}}=A^{\overline{\alpha}}_{\;\beta}x^\beta+\frac{1}{2}A^{\overline{\alpha}}_{\;\alpha}\Gamma^{\alpha}_{\;\beta\gamma}(p)x^\beta x^\gamma+O(x^3)
\ee
and the inverse transformation is just

\be\label{eq:RNCinversetrans}
x^{\alpha}=A^{\alpha}_{\;\overline{\beta}}y^{\overline{\beta}}+O(y^2)
\ee
One finds that the $A^{\alpha}_{\;\overline{\alpha}}$ also satisfy

\be\label{eq:RNCeqnforA}
A^{\overline{\alpha}}_{\;\mu}A^{\mu}_{\;\overline{\beta}}=\delta^{\overline{\alpha}}_{\;\overline{\beta}}\;,\;\;\;\;A^{\alpha}_{\;\overline{\mu}}A^{\overline{\mu}}_{\;\beta}=\delta^{\alpha}_{\;\beta}
\ee
These relations for RNC will cover all that will be needed in this discussion.

\section{Lightcone Volumes}

The volumes of the truncated lightcones can be found as follows. Consider the past lightcone emanating at the point $p_0=(T,\mathbf x_0)$. There is a unique point $q_0=(0,\mathbf x_0)$ on $\Sigma$ associated with $p_0$, separated from $p_0$ by a proper time $T$.

The volume of this cone is given by

\be\label{eq:VolumeWithNoSimplifications}
V(0,\mathbf{x}_0;T,\mathbf{x}_0)=\int_{\mathcal{X}} d^d x\;\sqrt{-g}
\ee
where the integration region, $\mathcal{X}:= J^-(p_0)\bigcap J^+(\Sigma)$ will be some complicated thing. As we are dealing with points close to the surface we can use the  Riemann normal coordinates $y^{\overline{\mu}}$, defined as above, about the point $q_0$ (the point $q_0$ being the origin of our RNC system). One finds that $A^0_{\;\overline{0}}=1$, $A^0_{\;i}=0$ and $\delta_{\overline{i}\overline{j}}=A^i_{\;\overline{i}}A^j_{\;\overline{j}}h_{ij}(0,\mathbf{x}_0)$. It can be shown that the volume integral reduces to \cite{Khetrapal_Sumati:Causal_Diamond_Volume}

\be\label{eq:VolumeWithRNC}
V(0,\mathbf{x}_0;T,\mathbf{x}_0) =\int_{\mathcal{X}}d^dy+\int_{\mathcal{X}_0}d^dy\left(-\frac{1}{6}R_{\overline{\mu}\overline{\nu}}(q_0)y^{\overline{\mu}}y^{\overline{\nu}} \right)+O(T^{d+3})
\ee
where $\mathcal{X}_0=\left\lbrace y^{\overline{\mu}} \mid 0\leq y^{\overline{0}}=\overline{t}\leq T\; ,\; \sqrt{(y^{\overline{1}})^2+(y^{\overline{2}})^2+...+(y^{\overline{d-1}})^2}\leq T-\overline{t} \right\rbrace$ is what we call the \textit{flat cone}, and $R_{\overline{\mu}\overline{\nu}}(q_0)$ is the Ricci tensor in RNC evaluated at $q_0$. The arguments of the volume function need not be changed as the volume only depends on the points in the manifold, $q_0$ and $p_0$. The \textit{flat cone} term comes in at $O(T^{d+2})$ so the volume we have to calculate has reduced to

\be\label{eq:VolumeToLowestOrder}
V(0,\mathbf{x}_0;T,\mathbf{x}_0) =\int_{\mathcal{X}}d^dy+O(T^{d+2})
\ee
Terms of $O(T^{d+2})$ can be retained till the end, but in the limit, $\rho\rightarrow\infty$, they vanish. This will be proved below.

The boundaries of $\mathcal{X}$ must now be figured out in order to do this integral. First we look at the top part of the cone. From \cite{Khetrapal_Sumati:Causal_Diamond_Volume} it can be shown that the first curvature correction to the top part of the cone comes in at $O(T^{d+2})$ and so can be ignored for our purposes. This means that the top can be treated as a that of a \textit{flat cone} and so points on the boundary obey the equation $\sqrt{(y^{\overline{1}})^2+(y^{\overline{2}})^2+...+(y^{\overline{d-1}})} = T-\overline{t}$.

The bottom surface of the cone was defined in terms of GNC to be $t=0$, so we can use (\ref{eq:RNCtotaltrans}) to find the equation for the surface in RNC. Equation (\ref{eq:RNCtotaltrans}) gives

\be\label{eq:BottomSurfaceWithGNC}
y^{\overline{0}}=\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)x^i x^j+O(x^3)
\ee
The first part on the right vanishes as $A^{\overline{0}}_{\;\mu}x^{\mu}=x^0$ (as $A^{\overline{0}}_{\;i}=0$ and $A^{\overline{0}}_{\;0}=1$) and $x^0=t=0$ for the bottom surface. Using the inverse RNC relation (\ref{eq:RNCinversetrans}) one can find the equation for the bottom surface entirely in RNC.

\be\label{eq:BottomSurface}
\overline{t}=\frac{1}{2}\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)A^{i}_{\;\overline{i}}A^{j}_{\;\overline{j}}y^{\overline{i}} y^{\overline{j}}+O(y^3)
\ee

This equation can be written in terms of the radial coordinate of the cone, $r=\sqrt{(y^{\overline{1}})^2+...+(y^{\overline{d-1}})^2}$, as

\be\label{eq:RadialBottomSurface}
\overline{t}=\frac{1}{2}\left(\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)A^{i}_{\;\overline{i}}A^{j}_{\;\overline{j}}\frac{y^{\overline{i}} y^{\overline{j}}}{r^2}\right)r^2+O(y^3)=\frac{1}{2}f(\mathbf{x}_0,\phi)r^2+O(y^3)
\ee
where $\phi$ stands for all the angular coordinates $\phi_1,..,\phi_{d-2}$. The function $f(\mathbf{x}_0,\phi)$ depends on $\mathbf{x}_0$ as $\Gamma^{0}_{\;ij}$ and $A^{i}_{\;\overline{i}}$ will, in general, be different for a different point on the surface. One can see that $f(\mathbf{x}_0,\phi)$ depends on the angles, $\phi$, by substituting the relations below into (\ref{eq:RadialBottomSurface})

\begin{align}\label{eq:SphericalCoords}
y^{\overline{1}} &= r \cos(\phi_1) \nonumber \\
y^{\overline{2}} &= r \sin(\phi_1) \cos(\phi_2) \nonumber \\
y^{\overline{2}} &= r \sin(\phi_1) \sin(\phi_2) \cos(\phi_3) \nonumber \\
    &\vdots \nonumber \\
y^{\overline{d-2}} &= r \sin(\phi_1) \cdots \sin(\phi_{d-3}) \cos(\phi_{d-2}) \nonumber \\
y^{\overline{d-1}} &= r \sin(\phi_1) \cdots \sin(\phi_{d-3}) \sin(\phi_{d-2})
\end{align}

With the boundaries of the integration region in place, we can now write down the integral explicitly, in spherical coordinates:

\be\label{eq:VolumeIntegralSpherical}
V(0,\mathbf{x}_0;T,\mathbf{x}_0)=\int_{S^{d-2}}
d\Omega_{d-2}
\int_{0}^{r_{max}(\phi)}r^{d-2}dr
\int_{\frac{1}{2}f(\mathbf{x}_0,\phi)r^2}^{-r+T}
d\overline{t}+O(T^{d+2})
\ee
where $r_{max}(\phi)$ is the radial coordinate of where the bottom surface meets the top part of the cone at an angle $\phi$, shown in \ref{fig:cone_plot}. To find this one needs to solve $\frac{1}{2}f(\mathbf{x}_0,\phi){r_{max}}^2(\phi)=-r_{max}(\phi)+T$ and take the positive solution. The solution can be expanded in $T$ and is simply $r_{max}=T+O(T^2)$, with angular dependent terms contributing at $O(T^2)$. The $O(T^2)$ term will contribute at $O(T^{d+2})$ in the integral and so can be ignored. Substituting $r_{max}=T$ into (\ref{eq:VolumeIntegralSpherical}) means it can be evaluated to give

\begin{figure}
  \centering
    \includegraphics[scale=0.6]{coneplot}
     \caption{The size of the region inside the top and bottom bounding surfaces is the volume we want to calculate.}
     \label{fig:cone_plot}
\end{figure}

\be\label{eq:VolumeNoK}
V(0,\mathbf{x}_0;T,\mathbf{x}_0)
=\frac{A_{d-2}}{d(d-1)}T^d\left(1-\frac{d}{2(d+1)}\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)A^{i}_{\;\overline{i}}A^{j}_{\;\overline{j}}\delta^{\overline{i}\overline{j}}T\right)
+O(T^{d+2})
\ee
where $A_{d-2}$ is the $d$-$2$ dimensional volume of a $d$-$2$ dimensional sphere and the $\delta^{\overline{i}\overline{j}}$ has come from the fact that cross terms ($\overline{i}\neq \overline{j}$) vanish under the angular integration. The defining relations for $A^{i}_{\;\overline{i}}$ can be rearranged to give $A^{i}_{\;\overline{i}}A^{j}_{\;\overline{j}}\delta^{\overline{i}\overline{j}}=h^{ij}(0,\mathbf{x}_0)$. In GNC it can be shown that the extrinsic curvature is given by

\be\label{eq:K}
K(0,\mathbf{x}_0)
=g^{\mu\nu }\nabla_{\mu}n_{\nu}
=-\Gamma^{0}_{\;ij}(0,\mathbf{x}_0)h^{ij}(0,\mathbf{x}_0)=-\frac{1}{2}\frac{\dot{h}(0,\mathbf{x}_0)}{h(0,\mathbf{x}_0)}
\ee
which means that the first correction to the volume of the cone depends on $K(0,\mathbf{x}_0)$, and thus the approximate volume of the cone is

\begin{align}
V(0,\mathbf{x}_0;T,\mathbf{x}_0)
&=\frac{A_{d-2}}{d(d-1)}T^d\left(1+\frac{d}{2(d+1)}K(0,\mathbf{x}_0)T\right)
+O(T^{d+2}) \label{eq:TopVolumeWithK}\\
V(-T,\mathbf{x}_0;0,\mathbf{x}_0)
&=\frac{A_{d-2}}{d(d-1)}T^d\left(1-\frac{d}{2(d+1)}K(0,\mathbf{x}_0)T\right)
+O(T^{d+2}) \label{eq:BottomVolumeWithK}
\end{align}
where (\ref{eq:BottomVolumeWithK}) is the volume of the cone with its tip below the surface, and can be found from similar arguments.

\section{Final Stretch of the Proof}

@@

\begin{gather}\label{eq:NmaxNminStart}
\begin{aligned}
\lim_{\rho \to \infty}\left\langle N_{max}-N_{min} \right\rangle &= \\
\lim_{\rho \to \infty}\rho
\int_{\Sigma}d^{d-1}x & \int_{-\varepsilon}^{0}dt\
h^{\frac{1}{2}}\left(1+
\frac{1}{2}\frac{\dot{h}}{h}t-\rho B(-1)^{d}t^{d+1}+\rho O(t^{d+2})\right)e^{-\rho A(-1)^{d}t^{d}} \\
-\rho\int_{\Sigma}d^{d-1}x &
\int_{0}^{\varepsilon}dt\
h^{\frac{1}{2}}\left(1+
\frac{1}{2}\frac{\dot{h}}{h}t-\rho Bt^{d+1}+\rho O(t^{d+2})\right)e^{-\rho At^{d}}
\end{aligned}
\end{gather}
where we have defined

\begin{gather}\label{A_and_B_defn}
\begin{aligned}
A & \equiv \frac{A_{d-2}}{d(d-1)} \\
B & \equiv \frac{A_{d-2}}{2(d-1)(d+1)}K(0,\mathbf{x})
\end{aligned}
\end{gather}
and we have Taylor expanded the $O(t^{d+1})$ parts of the volume functions, in small $t$, from the exponents. If we change to $u=-t$ in the first part on the right hand side, and $u=t$ on the second part then the integral can be simplified to

>>>>

% CURRENTLY BEING EDITED OFF-LINE BY: IAN

@@

\section{The main result for spacelike surfaces}
\begin{enumerate}
\item MB1: need to justify why higher order terms can be ignored.
\item MB1: why the threefold coordinate transformation? why no just RNCs at the base?
\item MB1: 
\end{enumerate}
Consider a sufficiently well-behaved spacetime $(M,g)$ and Cauchy surface $\Sigma$ in $M$. The causal past and future sets $M^\pm=J^\pm(\Sigma)$ form a partition of $M$ and $\partial M^\pm = \Sigma$. The Gibbons-Hawking-York boundary term for $M^\pm$ is in this case given by
$$
S_{GHY} = \pm\frac1{\kappa^2} \int_{\Sigma} d^{d-1}x \sqrt{-h} K
$$
where $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal $n$ \IGNORE{\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$}.

The extrinsic curvature of a spacelike hypersurface can be identified with the volume gradient across that hypersurface [[MB1: HOW?]]. On the causal set, the analogue of a spacelike hypersurface is a maximal anti-chain, i.e. a subset $A$ of $C$, no two elements of which are related. The intuitive analogue of the boundary term would then be the rate of change of the number of causal set elements below and above the antichain. We shall see that this intuitive idea indeed bears out.

Consider a causal set $\mathcal C$ obtained by a Poisson sprinkling of density $\rho$ into a spacetime $(M,g)$. A Cauchy surface $\Sigma\subset M$ induces a partition $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$ of the sprinkling, where $\mathcal C^\pm$ denotes the restriction of $\mathcal C$ to the points sprinkled to the causal future/past of $\Sigma$. Let us denote the number of maximal elements in $\mathcal C^-$ and the number of minimal elements in $\mathcal C^+$ by $N_{min}$ and $N_{max}$, respectively.
%\IGNORE{\be
%\begin{aligned}
%N_{min}=|\left\{x\in\mathcal C^+:\nexists\;y\in\mathcal C^+ s.t. \right\}|
%\end{aligned}
%\ee} 
We propose the following definition for the discrete Gibbons-Hawking-York boundary term for a causal set $\mathcal C$ in $d$ spacetime dimensions:
\be\label{GH_boundary_to_causet}
S^{(d)}_{GHY}[\mathcal C]=\rho^{\frac{2}{d}-1}c_{d}\left\langle N_{max}-N_{min}\right\rangle.
\ee
The constant $c_{d}$ only depends on the dimension and is given by

@@

To support this proposal we show that in the limit of infinite sprinkling density we obtain
\be
\lim_{\rho\rightarrow\infty}S^{(d)}_{GHY}[\mathcal C] = \frac1{\kappa^2} \int_{\Sigma} d^{d-1}x \sqrt{-h} K.\label{eq:mainconjecture}
\ee
%MB1: (too vague) During the calculation approximations will be made. These approximations are justified on the grounds that if certain orders are retained throughout the calculation they will in fact vanish in the limit $\rho \rightarrow \infty$

@@

[[IJ: NEED A FIGURE FOR THIS NEXT BIT]]

Both quantities will diverge, but if their difference $\langle N_{max}\rangle - \langle N_{min}\rangle = \langle N_{max} - N_{min}\rangle$ grows slower than or at order $\rho^{1-\frac2d}$, the proposed action~\eqref{GH_boundary_to_causet} will tend to a finite value in the continuum limit.

@@

\begin{align}\label{eq:nmax_and_eq:nmin}
\left\langle N_{max}\right\rangle & =\int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
h^{\frac{1}{2}}\left(1+
\frac{1}{2}\frac{\dot{h}}{h}t+O(t^2)\right)
 \rho\ e^{-\rho V(-t,\mathbf{x};0,\mathbf{x})}
\\
\left\langle N_{min}\right\rangle & =\int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
h^{\frac{1}{2}}\left(1+
\frac{1}{2}\frac{\dot{h}}{h}t+O(t^2)\right) \rho\ e^{-\rho V(0,\mathbf{x};t,\mathbf{x})}
\end{align}
where $h\equiv det\left(h_{ij}(0,\mathbf{x})\right)$ and $\dot{}\equiv \frac{\partial}{\partial t}$. The metric determinant has been in expanded in small $t$. The volume functions have been rewritten as follows: $V(x,\Sigma)=V(p,p_0)=V(-t,\mathbf{x};0,\mathbf{x})$ and $V(\Sigma,q)=V(q_0,q)=V(0,\mathbf{x};t,\mathbf{x})$

[[OLD SECTION THAT MB WANTED TO REWRITE]]
[[MB: Need to rewrite this paragraph]]. 
The volumes will depend on the proper time elapsed along the unique geodesic from a point in $\Sigma$ to a point $p$ such that the spatial coordinates remain unchanged. This is shown in Fig. . The coordinate time, $t$, is in fact the same as the proper time in the GNC we have chosen. Equations (\ref{eq:nmax}) and (\ref{eq:nmin}) can then be simplified as any time, $t$, which is far away from the surface will give rise to a large volume, which will then be exponentially suppressed as $\rho \rightarrow \infty$. We can choose a distance $\varepsilon$ away from the surface such that beyond it all volumes will be suppressed in the final limit. (\ref{eq:nmax}) and (\ref{eq:nmin}) are now given by

\begin{align}
\left\langle N_{max}\right\rangle & =\int_{\Sigma}d^{d-1}x\int_{-\varepsilon}^{0}dt\:
h^{\frac{1}{2}}\left(1+
\frac{1}{2}\frac{\dot{h}}{h}t+O(t^2)\right)
 \rho\ e^{-\rho V(t,\mathbf{x};0,\mathbf{x})}
\\
\left\langle N_{min}\right\rangle & =\int_{\Sigma}d^{d-1}x\int_{0}^{\varepsilon}dt\:
h^{\frac{1}{2}}\left(1+
\frac{1}{2}\frac{\dot{h}}{h}t+O(t^2)\right) \rho\ e^{-\rho V(0,\mathbf{x};t,\mathbf{x})}
\end{align}
where $h\equiv det\left(h_{ij}(0,\mathbf{x})\right)$ and $\dot{}\equiv \frac{\partial}{\partial t}$. The metric determinant has been in expanded in small $t$.
[[END OF OLD SECTION]]

The volumes can also be approximated using the fact that the tips of the cones are close to the surface, as $|t|\leq\varepsilon$ in the integral limits. The process for utilising the possible simplifications will be outlined below for the cone to the future of $\Sigma$, as the process for the past cone is nearly identical. Riemann normal coordinates are integral to the following calculation so it is worth noting the relevant formulas before jumping in.

The volumes of the truncated lightcones can be found as follows. Consider the past lightcone emanating at the point $p_0=(t_0,\mathbf x_0)$. There is a unique point $q_0=(0,\mathbf x_0)$ on $\Sigma$ associated with $p_0$, separated from $p_0$ by a proper time $t_0$. Now change  spatial coordinates to primed coordinates $(t',\mathbf{x} ')$, such that the spatial metric at $P$ is flat. This is equivalent to using Riemann normal coordinates (RNC), but just in the spatial dimensions. We require
\be\label{metric_flat_conditions}
h_{i'j'}(p_0)=\delta_{ij}  \;\;\text{and}\;\; \Gamma^{i'}_{j'k'}(p_0)=0.
\ee
This of course still allows for $\Gamma^{0}_{ij}(P)\neq0$. The surface $\Sigma$ still corresponds to $t'=t=0$ after this coordinate change. We now change from these coordinates to full spacetime RNC to calculate the volume. The new coordinates, $z^\mu$, are related to the previous coordinates via

\be\label{RNC_to_GNC_flat}
z^\mu=\delta^\mu_\nu y^\nu+\frac{1}{2}\delta^{\tilde{\mu}}_{\mu '}\Gamma^{\mu '}_{\alpha '\beta '}(P)x'^{\alpha '}x'^{\beta '}+O(x'^3)
\ee
This is the standard transformation to RNC. The inverse relations are given by

\be\label{inverse_trans}
x'^{\mu '}=\delta^{\mu '}_{\tilde{\mu}}\tilde{x}^{\tilde{\mu}}+O(\tilde{x}^2)
\ee

Using (\ref{inverse_trans}) in (\ref{RNC_to_GNC_flat})
 allows one to find the equation for the surface in RNC. It is given by

\be\label{RNC_surface}
\tilde{t}=\frac{1}{2}\Gamma^{t'}_{i'j'}(P)\delta^{i'}_{\tilde{i}}\delta^{j'}_{\tilde{j}}
\tilde{x}^{\tilde{i}}\tilde{x}^{\tilde{j}}+O(\tilde{x}^3)
\ee
when $t=t'=0$. It is a quadratic in the spatial variables.

Following (Sumati paper) we find the equation for the top part of the cone [[MB: need to define what you mean]] is simply

\be\label{top_of_cone_eqn}
\tilde{t}=-\tilde{r}+T
\ee
to the order we require. If we kept higher orders the top part would have an elliptic cross section. The volume would then acquire terms of $O(T^{n+2})$ which are suppressed in the final limit. The shape of the cone has been altered from the curvy cone to the one shown in Fig.

We now have our integral boundaries and we can attempt the integral. From (Sumati paper) the volume of the top cone can be written as [[MB1: integration region is wrong]]

\begin{gather}\label{volume_from_sumati}
\begin{aligned}
V(0,\mathbf{x}_0;T,\mathbf{x}_0) & =\int_{J^{+}(\Sigma)\bigcap J^{-}(T,\mathbf{\tilde{x}}_0)}d^n\tilde{x}\;  \sqrt{-g} \\
& =\int_{J^{+}(0,\mathbf{\tilde{x}})\bigcap J^{-}(T,\mathbf{\tilde{x}}_0)}d^n\tilde{x}\;\\
 & +\int_{flat\;cone}d^n\tilde{x}\left(-\frac{1}{6}\tilde{x}^{\tilde{c}}\tilde{x}^{\tilde{d}}
R_{\tilde{c}\tilde{d}}(0,\mathbf{\tilde{x}}_0)\right)
\end{aligned}
\end{gather}
where the flat cone is the cone you would have in a flat spacetime [[MB1: need to specify what you mean]]. The flat cone term comes in at $O(T^{n+2})$ and so can be ignored. The origins of the coordinate systems agree, as do the points at the tips of the cones in each coordinate system, thus the volume can be a function of the original GNC. The integral boundaries of the first term can be approximated from equations (\ref{RNC_surface}) and (\ref{top_of_cone_eqn}). Equation (\ref{top_of_cone_eqn}) can be written in the radial coordinate $\tilde{r}^2={\tilde{x}_{1}}^2+...+{\tilde{x}_{d-1}}^2$ as

\be\label{RNC_surface_in_r}
\tilde{t}=\frac{1}{2}\left(\Gamma^{t'}_{i'j'}(0,\mathbf{x}_0)\delta^{i'}_{\tilde{i}}
\delta^{j'}_{\tilde{j}}
\frac{\tilde{x}^{\tilde{i}}
\tilde{x}^{\tilde{j}}}{{\tilde{r}}^2}\right)
{\tilde{r}}^2
=\frac{1}{2}f(\tilde{\phi}){\tilde{r}}^2
\ee
where $f(\tilde{\phi})$ only depends on the angular coordinates $\tilde{\phi}_1,..,\tilde{\phi}_{n-2}$.

The volume then becomes

\be\label{volume_with_approx}
V(0,\mathbf{x}_0;T,\mathbf{x}_0)=\int_{S^{n-2}}
d\tilde{\Omega}_{n-2}
\int_{0}^{T}\tilde{r}^{n-2}dr
\int_{\frac{1}{2}f(\tilde{\phi})\tilde{r}^2}^{-\tilde{r}+T}
d\tilde{t}
\ee
where the max limit on $\tilde{r}$ comes from solving the equation $\frac{1}{2}f(\tilde{\phi})\tilde{r}^2_{max}=
-\tilde{r}_{max}+T$ and expanding in $T$ to the lowest order necessary. Evaluating this integral gives

\be\label{volume_no_K}
V(0,\mathbf{x}_0;T,\mathbf{x}_0)
=\frac{A_{n-2}}{n(d-1)}T^n\left(1-\frac{n}{2(n+1)}\Gamma^{t'}_{i'j'}(0,\mathbf{x}_0)\delta^{i'j'}T\right)
+O(T^{n+2})
\ee
We observe that $\Gamma^{t'}_{i'j'}(0,\mathbf{x}_0)\delta^{i'j'}$ is written in GNC with spatial flatness at $P$ (the $primed$ indices), although it is a scalar and so is independent of coordinate choice. The extrinsic curvature of the surface in these coordinates is given by

\be\label{K_eqn}
K(0,\mathbf{x}_0)
=g^{\mu '\nu '}\nabla_{\mu '}n_{\nu '}
=-\Gamma^{t'}_{i'j'}(0,\mathbf{x}_0)\delta^{i'j'}
\ee
where $n_{\nu '}=(1,0,0)$ is the surface normal. Thus we can directly substitute in for $K$ in (\ref{volume_no_K}) to obtain

\begin{align}
V(0,\mathbf{x}_0;T,\mathbf{x}_0)
&=\frac{A_{n-2}}{n(d-1)}T^n\left(1+\frac{n}{2(n+1)}K(0,\mathbf{x}_0)T\right)
+O(T^{n+2}) \label{top_volume_with_K}\\
V(-T,\mathbf{x}_0;0,\mathbf{x}_0)
&=\frac{A_{n-2}}{n(d-1)}T^n\left(1-\frac{n}{2(n+1)}K(0,\mathbf{x}_0)T\right)
+O(T^{n+2}) \label{bottom_volume_with_K}
\end{align}
where (\ref{bottom_volume_with_K}) is the volume for the bottom cone, and can be found from similar arguments. These results do not disagree with (Sumati paper or Myrhiem), which states that the first correction to a small causal diamond is $O(T^{n+2})$. The total volume for a causal diamond would be a sum of (\ref{top_volume_with_K}) and (\ref{bottom_volume_with_K}), and therefore $O(T^{n+1})$ terms would cancel.

@@

\begin{gather}\label{eq:nmax_eq:nmin_start}
\begin{aligned}
\lim_{\rho \to \infty}\left\langle N_{max}-N_{min} \right\rangle &= \\
\lim_{\rho \to \infty}\rho
\int_{\Sigma}d^{d-1}x & \int_{-\varepsilon}^{0}dt\
h^{\frac{1}{2}}\left(1+
\frac{1}{2}\frac{\dot{h}}{h}t-\rho B(-1)^{n}t^{n+1}\right)e^{-\rho A(-1)^{n}t^{n}} \\
-\rho\int_{\Sigma}d^{d-1}x &
\int_{0}^{\varepsilon}dt\
h^{\frac{1}{2}}\left(1+
\frac{1}{2}\frac{\dot{h}}{h}t-\rho Bt^{n+1}\right)e^{-\rho At^{n}}
\end{aligned}
\end{gather}
where we have defined

\begin{gather}\label{A_and_B_defn}
\begin{aligned}
A & \equiv \frac{A_{n-2}}{n(d-1)} \\
B & \equiv \frac{A_{n-2}}{2(d-1)(d+1)}K(0,\mathbf{x})
\end{aligned}
\end{gather}
and we have expanded the $O(t^{n+1})$ parts of the exponentials. This is allowed as the higher power of $t$ means they are suppressed more in the final limiting procedure. Using the fact that in GNCs, $K(0,\mathbf{x})=-\frac{1}{2}\frac{\dot{h}}{h}$, the integral can be found to be
